---
title: "Day 1: Data Wrangling & Visualization"
subtitle: "Afternoon Session | Data Analysis with R"
author: "IISc Bio Engineering Workshop"
date: "January 6, 2025"
format:
  revealjs:
    theme: [default, ../styles/custom.scss]
    slide-number: true
    code-line-numbers: false
    highlight-style: github
    transition: fade
---

## Afternoon Agenda

| Time | Topic |
|------|-------|
| 14:00 | dplyr: Data Wrangling |
| 15:00 | Exercises |
| 15:30 | Break |
| 15:45 | ggplot2: Visualization |
| 16:45 | Day 1 Practice |
| 17:00 | Wrap-up |

## The 5 dplyr Verbs

1. `filter()` - Pick rows
2. `select()` - Pick columns
3. `mutate()` - Create columns
4. `arrange()` - Sort rows
5. `summarize()` - Aggregate

## filter(): Pick Rows

```{r}
#| echo: true
library(tidyverse)

mtcars |>
  filter(mpg > 25)
```

## select(): Pick Columns

```{r}
#| echo: true
mtcars |>
  select(mpg, cyl, wt)
```

## mutate(): Create Columns

```{r}
#| echo: true
mtcars |>
  mutate(kpl = mpg * 0.425) |>
  select(mpg, kpl)
```

## arrange(): Sort Rows

```{r}
#| echo: true
mtcars |>
  arrange(desc(mpg)) |>
  head(3)
```

## summarize() + group_by()

```{r}
#| echo: true
mtcars |>
  group_by(cyl) |>
  summarize(
    mean_mpg = mean(mpg),
    n = n()
  )
```

## Chaining Operations

```{r}
#| echo: true
mtcars |>
  filter(hp > 100) |>
  group_by(cyl) |>
  summarize(mean_mpg = mean(mpg)) |>
  arrange(desc(mean_mpg))
```

## Exercise: dplyr Practice {.center}

Using `iris` dataset:

1. Filter for setosa species
2. Select Sepal columns only
3. Calculate mean Sepal.Length by Species

## ggplot2: Grammar of Graphics

```{r}
#| echo: true
#| output-location: column
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point()
```

## The ggplot Template

```r
ggplot(data, aes(x = ___, y = ___)) +
  geom_***()
```

- `data`: Your data frame
- `aes()`: Aesthetic mappings
- `geom_*()`: Geometric objects

## Common Geoms

| Geom | Use Case |
|------|----------|
| `geom_point()` | Scatter plot |
| `geom_line()` | Line plot |
| `geom_bar()` | Bar chart |
| `geom_boxplot()` | Box plot |
| `geom_histogram()` | Distribution |

## Adding Color

```{r}
#| echo: true
#| output-location: column
ggplot(mtcars, aes(x = wt, y = mpg,
                   color = factor(cyl))) +
  geom_point(size = 3)
```

## Faceting

```{r}
#| echo: true
#| fig-width: 10
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  facet_wrap(~cyl)
```

## Themes

```{r}
#| echo: true
#| output-location: column
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  theme_classic()
```

## Labels

```{r}
#| echo: true
#| output-location: column
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  labs(
    title = "Weight vs MPG",
    x = "Weight (1000 lbs)",
    y = "Miles per Gallon"
  ) +
  theme_classic()
```

## Exercise: ggplot2 {.center}

Create a scatter plot of `iris`:

- x = Sepal.Length
- y = Petal.Length
- color = Species
- Add labels and theme_classic()

## Day 1 Recap

✅ R fundamentals (vectors, data frames)

✅ Data import with readr

✅ The tidyverse & pipe operator

✅ dplyr: 5 core verbs

✅ ggplot2: Grammar of graphics

## Tomorrow: Day 2

- Descriptive statistics
- Hypothesis testing
- t-tests and ANOVA
- Correlation & regression
- Hands-on workshop

## Questions? {.center}

### See you tomorrow at 9:00 AM!
