---
title: "Day 2: Statistical Analysis"
subtitle: "Morning Session | Data Analysis with R"
author: "IISc Bio Engineering Workshop"
date: "January 7, 2025"
format:
  revealjs:
    theme: [default, ../styles/custom.scss]
    slide-number: true
    code-line-numbers: false
    highlight-style: github
    transition: fade
---

## Day 2 Agenda

| Time | Topic |
|------|-------|
| 09:00 | Descriptive Statistics |
| 10:00 | Hypothesis Testing Concepts |
| 10:30 | Break |
| 10:45 | t-tests |
| 12:00 | Exercises |
| 12:30 | Lunch |
| 14:00 | ANOVA |
| 15:00 | Correlation & Regression |
| 16:00 | Your Data Workshop |

## Descriptive Statistics

```{r}
#| echo: true
library(tidyverse)

mtcars |>
  summarize(
    mean_mpg = mean(mpg),
    sd_mpg = sd(mpg),
    median_mpg = median(mpg),
    n = n()
  )
```

## By Group

```{r}
#| echo: true
mtcars |>
  group_by(cyl) |>
  summarize(
    mean = mean(mpg),
    sd = sd(mpg),
    n = n()
  )
```

## Hypothesis Testing

### The Logic

1. State null hypothesis (H₀)
2. Collect data
3. Calculate test statistic
4. Determine p-value
5. Make decision

## P-value

**Definition**: Probability of observing data this extreme (or more) if H₀ is true

- p < 0.05 → "Statistically significant"
- p ≥ 0.05 → "Not significant"

⚠️ P-value ≠ probability that H₀ is true!

## Type I and Type II Errors

|  | H₀ True | H₀ False |
|--|---------|----------|
| Reject H₀ | Type I (α) | Correct |
| Accept H₀ | Correct | Type II (β) |

- **Type I**: False positive
- **Type II**: False negative

## The t-test

Compares means between groups

### Types:
1. One-sample t-test
2. Two-sample (independent) t-test
3. Paired t-test

## One-Sample t-test

*Is the mean different from a known value?*

```{r}
#| echo: true
t.test(mtcars$mpg, mu = 25)
```

## Two-Sample t-test

*Do two groups have different means?*

```{r}
#| echo: true
# Create example data
auto <- mtcars |> filter(am == 0) |> pull(mpg)
manual <- mtcars |> filter(am == 1) |> pull(mpg)

t.test(auto, manual)
```

## Formula Syntax

```{r}
#| echo: true
t.test(mpg ~ am, data = mtcars)
```

**Read as**: mpg *by* am (transmission type)

## Paired t-test

*Before/after measurements on same subjects*

```{r}
#| echo: true
# Simulated before/after data
set.seed(123)
before <- rnorm(20, 100, 15)
after <- before + rnorm(20, 10, 5)

t.test(before, after, paired = TRUE)
```

## Assumptions of t-test

1. **Independence**: Observations are independent
2. **Normality**: Data approximately normal
3. **Homogeneity**: Equal variances (for two-sample)

## Checking Normality

```{r}
#| echo: true
shapiro.test(mtcars$mpg)
```

p > 0.05 → Can assume normality

## Exercise {.center}

Test if mean Sepal.Length in `iris` differs from 5.5:

```{r}
#| echo: true
#| eval: false
t.test(iris$Sepal.Length, mu = 5.5)
```

## Break Time! {.center}

### 15 minutes

☕ See you back soon!
