---
title: "Day 2: ANOVA & Regression"
subtitle: "Afternoon Session | Data Analysis with R"
author: "IISc Bio Engineering Workshop"
date: "January 7, 2025"
format:
  revealjs:
    theme: [default, ../styles/custom.scss]
    slide-number: true
    code-line-numbers: false
    highlight-style: github
    transition: fade
---

## Afternoon Agenda

| Time | Topic |
|------|-------|
| 14:00 | ANOVA |
| 15:00 | Correlation |
| 15:30 | Break |
| 15:45 | Regression |
| 16:30 | Your Data Workshop |
| 17:00 | Wrap-up |

## ANOVA

**Analysis of Variance**

- Extension of t-test to 3+ groups
- Tests if means differ across groups
- H₀: All group means are equal

## One-Way ANOVA

```{r}
#| echo: true
library(tidyverse)

model <- aov(mpg ~ factor(cyl), data = mtcars)
summary(model)
```

## Interpretation

- **F value**: Ratio of between-group to within-group variance
- **Pr(>F)**: p-value
- If p < 0.05: At least one group differs

**But which groups?**

## Post-hoc: Tukey HSD

```{r}
#| echo: true
TukeyHSD(model)
```

## ANOVA Visualization

```{r}
#| echo: true
#| output-location: column
ggplot(mtcars, aes(factor(cyl), mpg,
                   fill = factor(cyl))) +
  geom_boxplot() +
  theme_classic() +
  theme(legend.position = "none")
```

## Correlation

**Measures linear relationship between two variables**

- r = 1: Perfect positive
- r = 0: No relationship
- r = -1: Perfect negative

## Calculating Correlation

```{r}
#| echo: true
cor(mtcars$wt, mtcars$mpg)
```

### With Hypothesis Test

```{r}
#| echo: true
cor.test(mtcars$wt, mtcars$mpg)
```

## Correlation Strength

| r | Strength |
|---|----------|
| 0.0 - 0.3 | Weak |
| 0.3 - 0.7 | Moderate |
| 0.7 - 1.0 | Strong |

## Visualizing Correlation

```{r}
#| echo: true
#| output-location: column
ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_classic()
```

## Simple Linear Regression

**Predicting one variable from another**

$$y = \beta_0 + \beta_1 x + \epsilon$$

- β₀: Intercept
- β₁: Slope
- ε: Error

## Fitting a Model

```{r}
#| echo: true
model <- lm(mpg ~ wt, data = mtcars)
summary(model)
```

## Interpretation

- **Intercept (37.29)**: Expected mpg when wt = 0
- **wt (-5.34)**: Each 1000 lb increase → 5.34 mpg decrease
- **R² (0.75)**: 75% of variance explained

## Multiple Regression

```{r}
#| echo: true
model2 <- lm(mpg ~ wt + hp + cyl, data = mtcars)
summary(model2)
```

## Tidy Output with broom

```{r}
#| echo: true
library(broom)
tidy(model)
```

## Model Statistics

```{r}
#| echo: true
glance(model)
```

## Your Data Workshop {.center}

### Practice Time!

Choose a dataset and apply what you've learned:

1. Descriptive statistics
2. Visualization
3. Statistical test
4. Interpretation

## Day 2 Recap

✅ Descriptive statistics

✅ Hypothesis testing concepts

✅ t-tests (one-sample, two-sample, paired)

✅ ANOVA and post-hoc tests

✅ Correlation and regression

## Tomorrow: Day 3

- Publication-quality figures
- Multi-panel figures with patchwork
- Case study analysis
- Quarto reports
- Mini project

## Questions? {.center}

### See you tomorrow at 9:00 AM!
