---
title: "Day 3: Publication Figures"
subtitle: "Morning Session | Data Analysis with R"
author: "IISc Bio Engineering Workshop"
date: "January 8, 2025"
format:
  revealjs:
    theme: [default, ../styles/custom.scss]
    slide-number: true
    code-line-numbers: false
    highlight-style: github
    transition: fade
---

## Day 3 Agenda

| Time | Topic |
|------|-------|
| 09:00 | Publication-Quality Figures |
| 10:00 | Multi-panel Figures |
| 10:30 | Break |
| 10:45 | Case Study |
| 12:30 | Lunch |
| 14:00 | Quarto Reports |
| 15:00 | Mini Project |
| 16:30 | Workshop Wrap-up |

## Publication Standards

### Common Requirements

- **Resolution**: 300+ DPI
- **Font size**: Minimum 8pt
- **Colors**: Colorblind-friendly
- **Format**: TIFF or PDF

## Custom Theme

```{r}
#| echo: true
library(tidyverse)

theme_publication <- function(base_size = 12) {
  theme_classic(base_size = base_size) +
    theme(
      plot.title = element_text(face = "bold"),
      axis.title = element_text(face = "bold"),
      panel.border = element_rect(fill = NA, color = "black"),
      legend.position = "bottom"
    )
}
```

## Colorblind-Friendly Palette

```{r}
#| echo: true
okabe_ito <- c("#E69F00", "#56B4E9", "#009E73",
               "#F0E442", "#0072B2", "#D55E00",
               "#CC79A7", "#999999")
```

## Applying the Theme

```{r}
#| echo: true
#| output-location: column
ggplot(mtcars, aes(wt, mpg,
                   color = factor(cyl))) +
  geom_point(size = 3) +
  scale_color_manual(values = okabe_ito) +
  labs(
    title = "Fuel Efficiency",
    x = "Weight (1000 lbs)",
    y = "MPG"
  ) +
  theme_publication()
```

## Exporting Figures

```{r}
#| echo: true
#| eval: false
# High resolution for journals
ggsave("figure1.tiff", width = 6, height = 4,
       dpi = 300, compression = "lzw")

# Vector for manuscripts
ggsave("figure1.pdf", width = 6, height = 4)

# For presentations
ggsave("figure1.png", width = 8, height = 6, dpi = 150)
```

## The patchwork Package

```{r}
#| echo: true
library(patchwork)
```

Combine ggplot2 plots easily!

## Creating Panels

```{r}
#| echo: true
p1 <- ggplot(mtcars, aes(wt, mpg)) +
  geom_point() + theme_classic()

p2 <- ggplot(mtcars, aes(factor(cyl), mpg)) +
  geom_boxplot() + theme_classic()

p3 <- ggplot(mtcars, aes(mpg)) +
  geom_histogram(bins = 10) + theme_classic()
```

## Side by Side

```{r}
#| echo: true
p1 + p2
```

## Stacked

```{r}
#| echo: true
p1 / p2
```

## Complex Layouts

```{r}
#| echo: true
(p1 + p2) / p3
```

## Adding Labels

```{r}
#| echo: true
(p1 + p2) / p3 +
  plot_annotation(tag_levels = "A")
```

## Shared Theme

```{r}
#| echo: true
(p1 + p2) / p3 &
  theme_classic(base_size = 10) &
  theme(plot.tag = element_text(face = "bold"))
```

## Collected Legends

```{r}
#| echo: true
#| fig-width: 10
pa <- ggplot(iris, aes(Sepal.Length, Petal.Length,
                       color = Species)) +
  geom_point() + theme_classic()

pb <- ggplot(iris, aes(Sepal.Width, Petal.Width,
                       color = Species)) +
  geom_point() + theme_classic()

pa + pb + plot_layout(guides = "collect")
```

## Exercise {.center}

Create a 2x2 figure with:

- Panel A: Scatter plot
- Panel B: Boxplot
- Panel C: Histogram
- Panel D: Bar chart

Add "A", "B", "C", "D" labels!

## Break Time! {.center}

### 15 minutes

☕ Next: Case Study Analysis
