---
title: "Day 2 Exercises: Statistical Analysis"
subtitle: "Instructor Version (Solutions)"
format: html
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(broom)
```

## Exercise 1: Descriptive Statistics

### 1.1 Basic Statistics

```{r}
iris |>
  summarize(
    mean_sl = mean(Sepal.Length),
    median_sl = median(Sepal.Length),
    sd_sl = sd(Sepal.Length)
  )
```

### 1.2 Group Statistics

```{r}
iris |>
  group_by(Species) |>
  summarize(
    n = n(),
    mean = mean(Sepal.Length),
    sd = sd(Sepal.Length)
  )
```

---

## Exercise 2: t-tests

### 2.1 One-Sample t-test

```{r}
t.test(iris$Sepal.Length, mu = 5.8)
```

**Interpretation**: p-value = 0.019 < 0.05, so we reject H0. The mean Sepal.Length is significantly different from 5.8.

### 2.2 Two-Sample t-test

```{r}
setosa <- iris |> filter(Species == "setosa") |> pull(Sepal.Length)
versicolor <- iris |> filter(Species == "versicolor") |> pull(Sepal.Length)

t.test(setosa, versicolor)
```

**Interpretation**: p < 2.2e-16, so there is a significant difference in Sepal.Length between setosa and versicolor.

### 2.3 Formula Syntax

```{r}
t.test(mpg ~ am, data = mtcars)
```

---

## Exercise 3: ANOVA

### 3.1 One-Way ANOVA

```{r}
model <- aov(Sepal.Length ~ Species, data = iris)
summary(model)
```

### 3.2 Post-hoc Test

```{r}
TukeyHSD(model)
```

**Interpretation**: All three pairwise comparisons are significant (all p adj < 0.05). Virginica has the longest sepals, setosa the shortest.

### 3.3 Visualization

```{r}
ggplot(iris, aes(x = Species, y = Sepal.Length, fill = Species)) +
  geom_boxplot(alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  labs(y = "Sepal Length (cm)") +
  theme_classic() +
  theme(legend.position = "none")
```

---

## Exercise 4: Correlation

### 4.1 Calculate Correlation

```{r}
cor(iris$Sepal.Length, iris$Petal.Length)
```

### 4.2 Correlation Test

```{r}
cor.test(iris$Sepal.Length, iris$Petal.Length)
```

**Interpretation**: r = 0.87, p < 2.2e-16. Strong positive correlation.

### 4.3 Scatter Plot

```{r}
ggplot(iris, aes(x = Sepal.Length, y = Petal.Length)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", color = "red") +
  labs(
    title = "Sepal vs Petal Length",
    x = "Sepal Length (cm)",
    y = "Petal Length (cm)"
  ) +
  theme_classic()
```

---

## Exercise 5: Regression

### 5.1 Simple Linear Regression

```{r}
model <- lm(Petal.Length ~ Sepal.Length, data = iris)
summary(model)
```

**Answers**:
- R-squared = 0.76 (76% of variance explained)
- Yes, significant (p < 2.2e-16)
- Slope = 1.86: For each 1 cm increase in Sepal.Length, Petal.Length increases by 1.86 cm

### 5.2 Multiple Regression

```{r}
model2 <- lm(mpg ~ wt + hp, data = mtcars)
summary(model2)
```

### 5.3 Tidy Output

```{r}
tidy(model2)
```

---

## Bonus Challenge

```{r}
# 1. Summary by supplement
ToothGrowth |>
  group_by(supp) |>
  summarize(
    n = n(),
    mean_len = mean(len),
    sd_len = sd(len)
  )

# 2. Boxplot
ggplot(ToothGrowth, aes(x = supp, y = len, fill = supp)) +
  geom_boxplot(alpha = 0.7) +
  labs(
    title = "Tooth Growth by Supplement Type",
    x = "Supplement",
    y = "Tooth Length"
  ) +
  theme_classic() +
  theme(legend.position = "none")

# 3. Statistical test
t.test(len ~ supp, data = ToothGrowth)

# 4. Interpretation: p = 0.06, marginally significant.
# OJ appears to have slightly higher tooth growth but
# the difference is not statistically significant at alpha = 0.05.
```
