---
title: "Day 2 Exercises: Statistical Analysis"
subtitle: "Participant Version"
format: html
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
```

## Exercise 1: Descriptive Statistics

Using the `iris` dataset:

### 1.1 Basic Statistics

Calculate mean, median, and SD for Sepal.Length:

```{r}
iris |>
  summarize(
    mean_sl = ___,
    median_sl = ___,
    sd_sl = ___
  )
```

### 1.2 Group Statistics

Calculate summary statistics for Sepal.Length BY Species:

```{r}
iris |>
  group_by(___) |>
  summarize(
    n = n(),
    mean = ___,
    sd = ___
  )
```

---

## Exercise 2: t-tests

### 2.1 One-Sample t-test

Test if mean Sepal.Length differs from 5.8:

```{r}
t.test(iris$Sepal.Length, mu = ___)
```

**Interpretation**: Is the p-value < 0.05? What do you conclude?

### 2.2 Two-Sample t-test

Compare Sepal.Length between setosa and versicolor:

```{r}
# Filter data
setosa <- iris |> filter(Species == "setosa") |> pull(Sepal.Length)
versicolor <- iris |> filter(Species == "versicolor") |> pull(Sepal.Length)

# Run t-test
t.test(___, ___)
```

### 2.3 Formula Syntax

Using mtcars, test if mpg differs by transmission type (am):

```{r}
t.test(___ ~ ___, data = ___)
```

---

## Exercise 3: ANOVA

### 3.1 One-Way ANOVA

Test if Sepal.Length differs across all 3 iris Species:

```{r}
model <- aov(___ ~ ___, data = ___)
summary(model)
```

### 3.2 Post-hoc Test

Run Tukey's HSD to find which species differ:

```{r}
TukeyHSD(___)
```

**Interpretation**: Which pairs of species are significantly different?

### 3.3 Visualization

Create a boxplot to visualize the differences:

```{r}
ggplot(iris, aes(x = ___, y = ___, fill = ___)) +
  geom_boxplot() +
  theme_classic()
```

---

## Exercise 4: Correlation

### 4.1 Calculate Correlation

Find the correlation between Sepal.Length and Petal.Length in iris:

```{r}
cor(iris$___, iris$___)
```

### 4.2 Correlation Test

Test if the correlation is significant:

```{r}
cor.test(___, ___)
```

### 4.3 Scatter Plot

Visualize the relationship:

```{r}
ggplot(iris, aes(x = Sepal.Length, y = Petal.Length)) +
  geom_point() +
  geom_smooth(method = "___") +
  theme_classic()
```

---

## Exercise 5: Regression

### 5.1 Simple Linear Regression

Predict Petal.Length from Sepal.Length:

```{r}
model <- lm(___ ~ ___, data = ___)
summary(model)
```

**Questions**:
- What is the R-squared value?
- Is the relationship significant?
- What does the slope mean?

### 5.2 Multiple Regression

Using mtcars, predict mpg from wt and hp:

```{r}
model2 <- lm(___ ~ ___ + ___, data = mtcars)
summary(model2)
```

### 5.3 Tidy Output

Use broom to get tidy coefficient table:

```{r}
library(broom)
tidy(___)
```

---

## Bonus Challenge

Perform a complete analysis on the `ToothGrowth` dataset:

1. Summarize tooth length by supplement type
2. Create a boxplot
3. Run an appropriate statistical test
4. Interpret your findings

```{r}
# Your code here
```
