---
title: "Day 1 Exercises: R Foundations"
subtitle: "Participant Version"
format: html
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
```

## Exercise 1: R Basics

### 1.1 Create Variables

Create the following variables:

```{r}
#| eval: false
# Patient age (45 years)
patient_age <- ___

# Drug dosage (2.5 mg)
drug_dose <- ___

# Sample ID ("SAMPLE_001")
sample_id <- ___
```

### 1.2 Create a Vector

Create a vector of 6 blood pressure readings: 120, 135, 118, 142, 128, 131

```{r}
#| eval: false
bp_readings <- ___

# Calculate the mean
mean_bp <- ___

# Calculate the standard deviation
sd_bp <- ___
```

### 1.3 Create a Data Frame

Create a data frame with 4 patients:

| ID | Age | Treatment |
|----|-----|-----------|
| P001 | 45 | Drug A |
| P002 | 32 | Placebo |
| P003 | 58 | Drug A |
| P004 | 41 | Placebo |

```{r}
#| eval: false
patients <- data.frame(
  id = ___,
  age = ___,
  treatment = ___
)
```

---

## Exercise 2: dplyr Practice

Use the `mtcars` dataset for these exercises.

### 2.1 Filter

Select only cars with more than 6 cylinders:

```{r}
#| eval: false
mtcars |>
  filter(___)
```

### 2.2 Select

Keep only the mpg, cyl, and hp columns:

```{r}
#| eval: false
mtcars |>
  select(___)
```

### 2.3 Mutate

Create a new column `kpl` (kilometers per liter) by multiplying mpg by 0.425:

```{r}
#| eval: false
mtcars |>
  mutate(kpl = ___) |>
  select(mpg, kpl)
```

### 2.4 Arrange

Sort by mpg in descending order:

```{r}
#| eval: false
mtcars |>
  arrange(___) |>
  head(5)
```

### 2.5 Summarize

Calculate mean and sd of mpg for each cylinder group:

```{r}
#| eval: false
mtcars |>
  group_by(___) |>
  summarize(
    mean_mpg = ___,
    sd_mpg = ___,
    n = ___
  )
```

---

## Exercise 3: ggplot2 Basics

### 3.1 Scatter Plot

Create a scatter plot of wt (x-axis) vs mpg (y-axis):

```{r}
#| eval: false
ggplot(mtcars, aes(x = ___, y = ___)) +
  ___()
```

### 3.2 Add Color

Color the points by number of cylinders:

```{r}
#| eval: false
ggplot(mtcars, aes(x = wt, y = mpg, color = ___)) +
  geom_point()
```

### 3.3 Add Labels

Add a title, x-label, and y-label:

```{r}
#| eval: false
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  labs(
    title = ___,
    x = ___,
    y = ___
  )
```

### 3.4 Box Plot

Create a boxplot of mpg by cylinder count:

```{r}
#| eval: false
ggplot(mtcars, aes(x = factor(cyl), y = ___)) +
  ___(fill = "steelblue") +
  theme_classic()
```

### 3.5 Complete Figure

Combine everything - create a scatter plot with:
- x = wt, y = mpg
- Color by cylinders
- Add a smooth line
- Add labels
- Use theme_classic()

```{r}
#| eval: false
ggplot(___, aes(___)) +
  geom_point(___) +
  geom_smooth(___) +
  labs(___) +
  theme_classic()
```

---

## Bonus Challenge

Using the `iris` dataset:

1. Filter for "setosa" species only
2. Calculate mean Sepal.Length and Petal.Length
3. Create a scatter plot of Sepal.Length vs Petal.Length

```{r}
# Your code here
```
