---
title: "Multi-Panel Figures"
subtitle: "Combining Plots with patchwork"
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(patchwork)
```

## Learning Objectives

1. Combine multiple plots into one figure
2. Use patchwork for layout control
3. Add panel labels (A, B, C...)

## The patchwork Package

```{r}
library(patchwork)
```

patchwork makes it easy to combine ggplot2 plots.

## Basic Combinations

### Creating Component Plots

```{r}
# Plot A: Scatter
p1 <- ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  labs(title = "Weight vs MPG") +
  theme_classic()

# Plot B: Boxplot
p2 <- ggplot(mtcars, aes(factor(cyl), mpg)) +
  geom_boxplot() +
  labs(title = "MPG by Cylinders", x = "Cylinders") +
  theme_classic()

# Plot C: Histogram
p3 <- ggplot(mtcars, aes(mpg)) +
  geom_histogram(bins = 10, fill = "steelblue") +
  labs(title = "MPG Distribution") +
  theme_classic()
```

### Side by Side

```{r}
p1 + p2
```

### Stacked

```{r}
p1 / p2
```

## Complex Layouts

### Three Panels

```{r}
(p1 + p2) / p3
```

### With Widths

```{r}
p1 + p2 + p3 + plot_layout(widths = c(2, 1, 1))
```

## Adding Panel Labels

```{r}
(p1 + p2) / p3 +
  plot_annotation(tag_levels = "A")
```

### Customize Labels

```{r}
(p1 + p2) / p3 +
  plot_annotation(
    tag_levels = "A",
    tag_prefix = "(",
    tag_suffix = ")"
  ) &
  theme(plot.tag = element_text(face = "bold", size = 14))
```

## Shared Elements

### Common Legend

```{r}
# Create plots with same color mapping
pa <- ggplot(iris, aes(Sepal.Length, Petal.Length, color = Species)) +
  geom_point() +
  theme_classic()

pb <- ggplot(iris, aes(Sepal.Width, Petal.Width, color = Species)) +
  geom_point() +
  theme_classic()

pa + pb + plot_layout(guides = "collect")
```

### Common Theme

```{r}
(p1 + p2) / p3 &
  theme_classic(base_size = 11) &
  theme(plot.title = element_text(size = 10, face = "bold"))
```

## Nested Layouts

```{r}
# Create fourth plot
p4 <- ggplot(mtcars, aes(factor(gear))) +
  geom_bar(fill = "coral") +
  labs(title = "Gear Count", x = "Gears") +
  theme_classic()

# Complex nesting
(p1 | (p2 / p4)) / p3 +
  plot_annotation(tag_levels = "A")
```

## Empty Spaces

```{r}
p1 + plot_spacer() + p2
```

## Inset Plots

```{r}
p1 + inset_element(p3, left = 0.6, bottom = 0.6, right = 1, top = 1)
```

## Complete Publication Figure

```{r}
#| fig-width: 10
#| fig-height: 8

# Theme for all
theme_pub <- theme_classic(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", size = 10),
    axis.title = element_text(size = 9),
    panel.border = element_rect(fill = NA, color = "black")
  )

# Recreate with consistent theme
fig_a <- ggplot(mtcars, aes(wt, mpg, color = factor(cyl))) +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_brewer(palette = "Set1", name = "Cylinders") +
  labs(title = "Fuel Efficiency", x = "Weight (1000 lbs)", y = "MPG") +
  theme_pub +
  theme(legend.position = c(0.8, 0.8))

fig_b <- ggplot(mtcars, aes(factor(cyl), mpg, fill = factor(cyl))) +
  geom_boxplot(show.legend = FALSE) +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Distribution by Cylinders", x = "Cylinders", y = "MPG") +
  theme_pub

fig_c <- ggplot(mtcars, aes(hp, mpg)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red") +
  labs(title = "Power vs Efficiency", x = "Horsepower", y = "MPG") +
  theme_pub

fig_d <- mtcars |>
  group_by(cyl) |>
  summarize(mean_mpg = mean(mpg), se = sd(mpg)/sqrt(n())) |>
  ggplot(aes(factor(cyl), mean_mpg, fill = factor(cyl))) +
  geom_col(show.legend = FALSE) +
  geom_errorbar(aes(ymin = mean_mpg - se, ymax = mean_mpg + se), width = 0.2) +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Mean MPG (± SE)", x = "Cylinders", y = "Mean MPG") +
  theme_pub

# Combine
(fig_a + fig_b) / (fig_c + fig_d) +
  plot_annotation(
    title = "Motor Vehicle Performance Analysis",
    subtitle = "Analysis of mtcars dataset (n = 32)",
    caption = "Figure 1. Relationship between vehicle characteristics and fuel efficiency.",
    tag_levels = "A",
    theme = theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 11, color = "gray40"),
      plot.caption = element_text(hjust = 0, size = 9)
    )
  )
```

## Exercise

::: {.callout-caution title="Try It"}
Create a 2x2 figure with `iris` data:

- Panel A: Sepal.Length vs Sepal.Width scatter
- Panel B: Petal.Length vs Petal.Width scatter
- Panel C: Sepal.Length boxplot by Species
- Panel D: Petal.Length boxplot by Species

Add "A", "B", "C", "D" labels.
:::

::: {.callout-tip collapse="true" title="Solution"}
```{r}
pa <- ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) +
  geom_point() + theme_classic() + theme(legend.position = "none")

pb <- ggplot(iris, aes(Petal.Length, Petal.Width, color = Species)) +
  geom_point() + theme_classic()

pc <- ggplot(iris, aes(Species, Sepal.Length, fill = Species)) +
  geom_boxplot() + theme_classic() + theme(legend.position = "none")

pd <- ggplot(iris, aes(Species, Petal.Length, fill = Species)) +
  geom_boxplot() + theme_classic() + theme(legend.position = "none")

(pa + pb) / (pc + pd) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A")
```
:::

---

**Next:** [Case Study: Biomedical Analysis](03_case_study_dfu.qmd)
