---
title: "Publication-Quality Figures"
subtitle: "Professional Visualization with ggplot2"
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(patchwork)
library(scales)
```

## Learning Objectives

1. Create publication-ready figures
2. Apply consistent styling
3. Export figures in appropriate formats

## Publication Standards

### Common Requirements

- **Resolution**: 300+ DPI for print
- **Font size**: Minimum 8pt in final figure
- **Colors**: Colorblind-friendly palettes
- **Format**: TIFF/PDF for journals

## Professional Theme

```{r}
theme_publication <- function(base_size = 12) {
  theme_classic(base_size = base_size) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0),
      plot.subtitle = element_text(color = "gray40"),
      axis.title = element_text(face = "bold"),
      axis.text = element_text(color = "black"),
      legend.position = "bottom",
      legend.title = element_text(face = "bold"),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 0.5),
      strip.background = element_rect(fill = "gray95", color = "black"),
      strip.text = element_text(face = "bold")
    )
}
```

## Colorblind-Friendly Palettes

```{r}
# Okabe-Ito palette
okabe_ito <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442",
               "#0072B2", "#D55E00", "#CC79A7", "#999999")

# Example usage
ggplot(iris, aes(x = Sepal.Length, y = Petal.Length, color = Species)) +
  geom_point(size = 2) +
  scale_color_manual(values = okabe_ito) +
  labs(title = "Iris Measurements", x = "Sepal Length (cm)", y = "Petal Length (cm)") +
  theme_publication()
```

## Adding Statistical Annotations

```{r}
#| eval: false
# Create sample data
set.seed(2025)
treatment_data <- tibble(
  group = rep(c("Control", "Treatment"), each = 30),
  value = c(rnorm(30, 50, 10), rnorm(30, 65, 12))
)

# Bar plot with error bars
summary_data <- treatment_data |>
  group_by(group) |>
  summarize(
    mean = mean(value),
    se = sd(value) / sqrt(n()),
    .groups = "drop"
  )

ggplot(summary_data, aes(x = group, y = mean, fill = group)) +
  geom_col(width = 0.6) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2) +
  geom_signif(comparisons = list(c("Control", "Treatment")),
              annotations = "***", y_position = 85) +
  scale_fill_manual(values = okabe_ito[c(1, 3)]) +
  labs(y = "Value (mean ± SE)") +
  theme_publication() +
  theme(legend.position = "none")
```

Note: `geom_signif()` requires the `ggsignif` package.

## Exporting Figures

### Using ggsave

```{r}
#| eval: false
# For journals (high resolution)
ggsave("figure1.tiff", width = 6, height = 4, dpi = 300, compression = "lzw")

# For manuscripts (vector)
ggsave("figure1.pdf", width = 6, height = 4)

# For presentations
ggsave("figure1.png", width = 8, height = 6, dpi = 150)
```

### Size Guidelines

| Journal | Single Column | Double Column |
|---------|---------------|---------------|
| Nature | 89 mm | 183 mm |
| Science | 85 mm | 176 mm |
| PLOS | 83 mm | 173 mm |

## Polished Example

```{r}
# Complete publication figure
p <- ggplot(mtcars, aes(x = wt, y = mpg, color = factor(cyl))) +
  geom_point(size = 3, alpha = 0.8) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1) +
  scale_color_manual(
    values = okabe_ito[c(1, 3, 5)],
    name = "Cylinders"
  ) +
  labs(
    title = "Fuel Efficiency vs Vehicle Weight",
    subtitle = "Linear regression by cylinder count",
    x = "Weight (1000 lbs)",
    y = "Fuel Efficiency (mpg)",
    caption = "Data: mtcars | n = 32"
  ) +
  theme_publication() +
  theme(legend.position = c(0.85, 0.85))

p
```

## Exercise

::: {.callout-caution title="Try It"}
Create a publication-quality boxplot of `iris` Sepal.Length by Species:

1. Use the Okabe-Ito palette
2. Apply `theme_publication()`
3. Add proper axis labels with units
4. Add a title and caption
:::

::: {.callout-tip collapse="true" title="Solution"}
```{r}
ggplot(iris, aes(x = Species, y = Sepal.Length, fill = Species)) +
  geom_boxplot(alpha = 0.8, outlier.shape = 21) +
  scale_fill_manual(values = okabe_ito[1:3]) +
  labs(
    title = "Sepal Length Distribution by Species",
    x = "Iris Species",
    y = "Sepal Length (cm)",
    caption = "Data: Anderson (1935) | n = 150"
  ) +
  theme_publication() +
  theme(legend.position = "none")
```
:::

---

**Next:** [Multi-Panel Figures](02_multipanel_figures.qmd)
