---
title: "Case Study: Biomedical Analysis"
subtitle: "Complete Analysis Workflow"
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(patchwork)
library(broom)
```

## Learning Objectives

1. Apply complete analysis workflow
2. Integrate data wrangling, statistics, and visualization
3. Practice biomedical data interpretation

## The Dataset: Cell Viability Study

Simulated study of drug treatment effects on cell viability.

```{r}
set.seed(2025)

# Generate synthetic biomedical data
cell_study <- tibble(
  sample_id = paste0("S", sprintf("%03d", 1:120)),
  treatment = rep(c("Control", "Drug_A", "Drug_B", "Combo"), each = 30),
  cell_line = rep(rep(c("HeLa", "MCF7", "A549"), each = 10), 4),
  viability = c(
    rnorm(30, 95, 5),   # Control
    rnorm(30, 75, 8),   # Drug A
    rnorm(30, 70, 10),  # Drug B
    rnorm(30, 45, 12)   # Combo
  ) |> pmax(0) |> pmin(100),  # Bound 0-100%
  apoptosis = 100 - viability + rnorm(120, 0, 5) |> pmax(0),
  concentration_um = rep(c(0, 10, 25, 50), each = 30)
)

glimpse(cell_study)
```

## Step 1: Exploratory Analysis

### Summary Statistics

```{r}
cell_study |>
  group_by(treatment) |>
  summarize(
    n = n(),
    mean_viability = mean(viability),
    sd_viability = sd(viability),
    mean_apoptosis = mean(apoptosis),
    .groups = "drop"
  ) |>
  arrange(mean_viability)
```

### By Cell Line

```{r}
cell_study |>
  group_by(treatment, cell_line) |>
  summarize(
    mean_viability = mean(viability),
    se = sd(viability) / sqrt(n()),
    .groups = "drop"
  )
```

## Step 2: Visualization

### Distribution Overview

```{r}
ggplot(cell_study, aes(x = treatment, y = viability, fill = treatment)) +
  geom_boxplot(alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5, size = 1) +
  scale_fill_viridis_d(option = "plasma", end = 0.8) +
  labs(
    title = "Cell Viability by Treatment",
    x = "Treatment Group",
    y = "Viability (%)"
  ) +
  theme_classic() +
  theme(legend.position = "none")
```

### By Cell Line

```{r}
ggplot(cell_study, aes(x = treatment, y = viability, fill = cell_line)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_brewer(palette = "Set2", name = "Cell Line") +
  labs(
    title = "Cell Viability by Treatment and Cell Line",
    x = "Treatment Group",
    y = "Viability (%)"
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Viability vs Apoptosis

```{r}
ggplot(cell_study, aes(x = viability, y = apoptosis, color = treatment)) +
  geom_point(alpha = 0.7, size = 2) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  labs(
    title = "Viability vs Apoptosis by Treatment",
    x = "Cell Viability (%)",
    y = "Apoptosis (%)"
  ) +
  theme_classic()
```

## Step 3: Statistical Analysis

### ANOVA: Treatment Effect

```{r}
model_treatment <- aov(viability ~ treatment, data = cell_study)
summary(model_treatment)
```

### Post-hoc Comparisons

```{r}
TukeyHSD(model_treatment)
```

### Two-way ANOVA: Treatment × Cell Line

```{r}
model_2way <- aov(viability ~ treatment * cell_line, data = cell_study)
summary(model_2way)
```

### Correlation Analysis

```{r}
cor.test(cell_study$viability, cell_study$apoptosis)
```

### Regression: Dose-Response

```{r}
model_dose <- lm(viability ~ concentration_um, data = cell_study)
tidy(model_dose)
glance(model_dose)
```

## Step 4: Publication Figure

```{r}
#| fig-width: 10
#| fig-height: 8

# Theme
theme_pub <- theme_classic(base_size = 10) +
  theme(
    plot.title = element_text(face = "bold", size = 11),
    axis.title = element_text(face = "bold"),
    panel.border = element_rect(fill = NA, color = "black")
  )

# Panel A: Viability by treatment
fig_a <- cell_study |>
  group_by(treatment) |>
  summarize(
    mean = mean(viability),
    se = sd(viability) / sqrt(n())
  ) |>
  ggplot(aes(x = reorder(treatment, -mean), y = mean, fill = treatment)) +
  geom_col(width = 0.7) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2) +
  scale_fill_viridis_d(option = "plasma", end = 0.8) +
  labs(title = "Cell Viability by Treatment", x = NULL, y = "Viability (%)") +
  theme_pub +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))

# Panel B: By cell line
fig_b <- ggplot(cell_study, aes(x = treatment, y = viability, fill = cell_line)) +
  geom_boxplot(alpha = 0.8) +
  scale_fill_brewer(palette = "Set2", name = "Cell Line") +
  labs(title = "Treatment Response by Cell Line", x = NULL, y = "Viability (%)") +
  theme_pub +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = c(0.85, 0.85))

# Panel C: Dose response
fig_c <- ggplot(cell_study, aes(x = concentration_um, y = viability)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "red") +
  labs(title = "Dose-Response Relationship",
       x = "Concentration (μM)", y = "Viability (%)") +
  theme_pub

# Panel D: Viability vs Apoptosis
fig_d <- ggplot(cell_study, aes(x = viability, y = apoptosis, color = treatment)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.8) +
  scale_color_viridis_d(option = "plasma", end = 0.8, name = "Treatment") +
  labs(title = "Viability-Apoptosis Correlation",
       x = "Viability (%)", y = "Apoptosis (%)") +
  theme_pub +
  theme(legend.position = "bottom")

# Combine
(fig_a + fig_b) / (fig_c + fig_d) +
  plot_annotation(
    title = "Drug Treatment Effects on Cell Viability",
    subtitle = "In vitro analysis across three cell lines (n = 120)",
    caption = "Figure 2. Treatment significantly reduced cell viability (p < 0.001). Combination therapy showed greatest effect.",
    tag_levels = "A"
  ) &
  theme(
    plot.tag = element_text(face = "bold", size = 12)
  )
```

## Step 5: Summary Table

```{r}
cell_study |>
  group_by(treatment) |>
  summarize(
    n = n(),
    `Viability (mean)` = round(mean(viability), 1),
    `Viability (SD)` = round(sd(viability), 1),
    `Apoptosis (mean)` = round(mean(apoptosis), 1),
    `Apoptosis (SD)` = round(sd(apoptosis), 1)
  ) |>
  knitr::kable()
```

## Key Findings

1. **Treatment Effect**: Significant difference in viability across treatments (F = X, p < 0.001)
2. **Combination Therapy**: Combo treatment showed lowest viability (~45%)
3. **Cell Line**: No significant interaction with cell line
4. **Correlation**: Strong negative correlation between viability and apoptosis (r = -0.9)

## Exercise

::: {.callout-caution title="Try It"}
Analyze the relationship between `concentration_um` and `apoptosis`:

1. Create a scatter plot with regression line
2. Calculate correlation
3. Fit a linear model and interpret
:::

::: {.callout-tip collapse="true" title="Solution"}
```{r}
# Scatter plot
ggplot(cell_study, aes(x = concentration_um, y = apoptosis)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "darkred") +
  labs(x = "Concentration (μM)", y = "Apoptosis (%)") +
  theme_classic()

# Correlation
cor.test(cell_study$concentration_um, cell_study$apoptosis)

# Linear model
model <- lm(apoptosis ~ concentration_um, data = cell_study)
summary(model)
```
:::

---

**Next:** [Quarto Reports](04_quarto_reports.qmd)
