---
title: "Mini Project"
subtitle: "Putting It All Together"
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(patchwork)
```

## Learning Objectives

1. Complete an end-to-end analysis
2. Create a reproducible report
3. Practice all workshop skills

## Your Task

Create a complete analysis report using a dataset of your choice.

## Dataset Options

### Option 1: Built-in Datasets

```{r}
#| eval: false
# Explore available datasets
data()

# Good choices:
# - mtcars: Vehicle performance
# - iris: Flower measurements
# - ToothGrowth: Guinea pig teeth
# - PlantGrowth: Plant experiment
# - ChickWeight: Chick growth
```

### Option 2: Sample Biomedical Data

```{r}
# Generate sample clinical trial data
set.seed(2025)

clinical_trial <- tibble(
  patient_id = paste0("PT", sprintf("%03d", 1:100)),
  age = round(rnorm(100, 55, 12)),
  sex = sample(c("M", "F"), 100, replace = TRUE),
  treatment = sample(c("Placebo", "Low Dose", "High Dose"), 100, replace = TRUE),
  baseline_score = round(rnorm(100, 50, 15)),
  final_score = baseline_score + case_when(
    treatment == "Placebo" ~ rnorm(100, 2, 8),
    treatment == "Low Dose" ~ rnorm(100, 10, 10),
    treatment == "High Dose" ~ rnorm(100, 18, 12)
  ) |> round(),
  adverse_event = sample(c(TRUE, FALSE), 100, replace = TRUE, prob = c(0.2, 0.8))
) |>
  mutate(
    change = final_score - baseline_score,
    responder = change > 10
  )

glimpse(clinical_trial)
```

### Option 3: Your Own Data

Bring a CSV file and use:

```{r}
#| eval: false
my_data <- read_csv("path/to/your/data.csv")
```

## Report Requirements

Your report should include:

### 1. Introduction

- Brief description of the dataset
- Research question(s)

### 2. Methods

- Data source
- Statistical approaches used

### 3. Results

#### Descriptive Statistics

```{r}
clinical_trial |>
  group_by(treatment) |>
  summarize(
    n = n(),
    mean_age = mean(age),
    pct_female = mean(sex == "F") * 100,
    mean_change = mean(change),
    pct_responder = mean(responder) * 100
  ) |>
  knitr::kable(digits = 1)
```

#### Visualization (minimum 2 figures)

```{r}
# Figure 1: Distribution
p1 <- ggplot(clinical_trial, aes(x = treatment, y = change, fill = treatment)) +
  geom_boxplot(alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  labs(title = "Score Change by Treatment", y = "Change from Baseline") +
  theme_classic() +
  theme(legend.position = "none")

# Figure 2: Response rate
response_summary <- clinical_trial |>
  group_by(treatment) |>
  summarize(response_rate = mean(responder) * 100)

p2 <- ggplot(response_summary, aes(x = treatment, y = response_rate, fill = treatment)) +
  geom_col() +
  geom_text(aes(label = paste0(round(response_rate), "%")), vjust = -0.5) +
  labs(title = "Response Rate by Treatment", y = "Responders (%)") +
  ylim(0, 100) +
  theme_classic() +
  theme(legend.position = "none")

p1 + p2
```

#### Statistical Test

```{r}
# ANOVA
model <- aov(change ~ treatment, data = clinical_trial)
summary(model)

# Post-hoc
TukeyHSD(model)
```

### 4. Discussion

- Key findings
- Limitations
- Conclusions

## Report Template

Create a new file called `my_analysis.qmd`:

````markdown
---
title: "Your Analysis Title"
author: "Your Name"
date: today
format:
  html:
    toc: true
    code-fold: true
execute:
  warning: false
  message: false
---

## Introduction

[Describe your data and research question]

## Methods

```{{r}}
#| label: setup
library(tidyverse)
library(patchwork)
```

```{{r}}
#| label: load-data
# Load or create your data
```

## Results

### Descriptive Statistics

```{{r}}
#| label: tbl-summary
#| tbl-cap: "Summary statistics"
# Your summary table
```

### Visualization

```{{r}}
#| label: fig-main
#| fig-cap: "Main results figure"
#| fig-width: 10
#| fig-height: 6
# Your multi-panel figure
```

### Statistical Analysis

```{{r}}
#| label: stats
# Your statistical tests
```

## Discussion

[Interpret your findings]

## References

[If applicable]
````

## Checklist

Before submitting, verify:

- [ ] Document renders without errors
- [ ] All figures have captions
- [ ] All tables are formatted
- [ ] Code is clean and commented
- [ ] Results are interpreted
- [ ] Reproducible (runs from fresh session)

## Sharing Your Work

### Export Options

```{r}
#| eval: false
# Render to HTML
quarto::quarto_render("my_analysis.qmd")

# Render to PDF
quarto::quarto_render("my_analysis.qmd", output_format = "pdf")
```

### Save Key Figures

```{r}
#| eval: false
# Save publication figure
ggsave("figures/main_figure.png", width = 10, height = 6, dpi = 300)
ggsave("figures/main_figure.pdf", width = 10, height = 6)
```

## Example Complete Analysis

::: {.callout-tip collapse="true" title="Full Example"}
```{r}
#| fig-width: 10
#| fig-height: 8

# Complete analysis of clinical trial data

# 1. Summary statistics
summary_stats <- clinical_trial |>
  group_by(treatment) |>
  summarize(
    n = n(),
    `Age (mean ± SD)` = paste0(round(mean(age), 1), " ± ", round(sd(age), 1)),
    `% Female` = round(mean(sex == "F") * 100, 1),
    `Baseline Score` = round(mean(baseline_score), 1),
    `Final Score` = round(mean(final_score), 1),
    `Change` = round(mean(change), 1),
    `% Responders` = round(mean(responder) * 100, 1)
  )

knitr::kable(summary_stats, caption = "Patient Characteristics by Treatment Group")

# 2. Multi-panel figure
fig_a <- ggplot(clinical_trial, aes(x = baseline_score, y = final_score, color = treatment)) +
  geom_point(alpha = 0.7) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  labs(title = "Baseline vs Final Score", x = "Baseline", y = "Final") +
  theme_classic()

fig_b <- ggplot(clinical_trial, aes(x = treatment, y = change, fill = treatment)) +
  geom_boxplot(alpha = 0.7) +
  geom_hline(yintercept = 10, linetype = "dashed", color = "red") +
  labs(title = "Change from Baseline", x = NULL, y = "Change") +
  theme_classic() +
  theme(legend.position = "none")

fig_c <- clinical_trial |>
  group_by(treatment) |>
  summarize(rate = mean(responder) * 100) |>
  ggplot(aes(x = treatment, y = rate, fill = treatment)) +
  geom_col() +
  labs(title = "Response Rate", x = NULL, y = "Responders (%)") +
  theme_classic() +
  theme(legend.position = "none")

fig_d <- ggplot(clinical_trial, aes(x = age, y = change, color = treatment)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Age vs Treatment Response", x = "Age", y = "Change") +
  theme_classic()

(fig_a + fig_b) / (fig_c + fig_d) +
  plot_layout(guides = "collect") +
  plot_annotation(
    title = "Clinical Trial Results",
    tag_levels = "A"
  ) &
  theme(legend.position = "bottom")

# 3. Statistical test
cat("\n--- ANOVA Results ---\n")
model <- aov(change ~ treatment, data = clinical_trial)
print(summary(model))
```
:::

---

**Workshop Complete!**

Thank you for participating. Continue practicing with your own data!
