---
title: "Quarto Reports"
subtitle: "Reproducible Documents"
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
```

## Learning Objectives

1. Understand Quarto document structure
2. Use code chunks effectively
3. Create reproducible reports

## What is Quarto?

- Next-generation R Markdown
- Supports R, Python, Julia, Observable
- Multiple output formats (HTML, PDF, Word, slides)
- Used for this entire workshop!

## Document Structure

```markdown
---
title: "My Analysis"
author: "Your Name"
date: today
format: html
---

## Introduction

Text goes here.
```

## Code Chunks

### Basic Chunk

````markdown
```{{r}}
summary(mtcars)
```
````

### Chunk Options

````markdown
```{{r}}
#| label: my-plot
#| fig-width: 8
#| fig-height: 6
#| fig-cap: "Scatter plot of weight vs mpg"

ggplot(mtcars, aes(wt, mpg)) +
  geom_point()
```
````

## Common Chunk Options

| Option | Purpose | Example |
|--------|---------|---------|
| `label` | Name the chunk | `#| label: setup` |
| `echo` | Show code | `#| echo: false` |
| `eval` | Run code | `#| eval: false` |
| `include` | Show anything | `#| include: false` |
| `warning` | Show warnings | `#| warning: false` |
| `message` | Show messages | `#| message: false` |
| `fig-width` | Figure width | `#| fig-width: 8` |
| `fig-cap` | Figure caption | `#| fig-cap: "My figure"` |

## Inline Code

Use backticks with `r` for inline calculations:

```markdown
The mean mpg is `r knitr::inline_expr("mean(mtcars$mpg)")`.
```

Renders as: The mean mpg is `r round(mean(mtcars$mpg), 1)`.

## Cross-References

### Figures

````markdown
```{{r}}
#| label: fig-scatter
#| fig-cap: "MPG vs Weight"

ggplot(mtcars, aes(wt, mpg)) + geom_point()
```

See @fig-scatter for the relationship.
````

### Tables

````markdown
```{{r}}
#| label: tbl-summary
#| tbl-cap: "Summary statistics"

mtcars |> summarize(mean_mpg = mean(mpg)) |> knitr::kable()
```

See @tbl-summary for details.
````

## Callout Boxes

::: {.callout-note}
This is a note callout.
:::

::: {.callout-tip}
This is a tip callout.
:::

::: {.callout-warning}
This is a warning callout.
:::

::: {.callout-caution}
This is a caution callout.
:::

```markdown
::: {.callout-note}
This is a note callout.
:::
```

## Collapsible Content

::: {.callout-tip collapse="true" title="Click to expand"}
This content is hidden by default!
:::

```markdown
::: {.callout-tip collapse="true" title="Click to expand"}
Hidden content here.
:::
```

## Tabbed Sections

::: {.panel-tabset}

### R Code

```{r}
mean(mtcars$mpg)
```

### Python Equivalent

```python
import pandas as pd
mtcars = pd.read_csv("mtcars.csv")
mtcars["mpg"].mean()
```

:::

````markdown
::: {.panel-tabset}

### R Code
```{{r}}
mean(mtcars$mpg)
```

### Python
```{{python}}
# Python code
```

:::
````

## Output Formats

### HTML (default)

```yaml
format: html
```

### PDF (requires LaTeX)

```yaml
format: pdf
```

### Word

```yaml
format: docx
```

### Multiple Formats

```yaml
format:
  html: default
  pdf: default
  docx: default
```

## Report Template

```yaml
---
title: "Analysis Report"
author: "Your Name"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    theme: cosmo
execute:
  warning: false
  message: false
---
```

## Rendering

From terminal:
```bash
quarto render report.qmd
```

From R:
```r
quarto::quarto_render("report.qmd")
```

In RStudio: Click "Render" button or Ctrl/Cmd + Shift + K

## Best Practices

1. **Name all chunks** with descriptive labels
2. **Set global options** in YAML header
3. **Use `here::here()`** for file paths
4. **Load packages** in setup chunk
5. **Set seed** for reproducibility
6. **Cache** long computations

## Exercise

::: {.callout-caution title="Try It"}
Create a simple Quarto document that:

1. Loads the iris dataset
2. Shows a summary table
3. Creates a figure with caption
4. Uses at least one callout box
:::

::: {.callout-tip collapse="true" title="Solution"}
````markdown
---
title: "Iris Analysis"
format: html
---

## Data Overview

```{{r}}
#| label: setup
library(tidyverse)
```

::: {.callout-note}
The iris dataset contains 150 observations of 3 species.
:::

```{{r}}
#| label: tbl-summary
#| tbl-cap: "Iris summary by species"
iris |>
  group_by(Species) |>
  summarize(across(everything(), mean)) |>
  knitr::kable(digits = 2)
```

```{{r}}
#| label: fig-iris
#| fig-cap: "Sepal dimensions by species"
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) +
  geom_point() +
  theme_classic()
```

See @fig-iris for the scatter plot and @tbl-summary for means.
````
:::

---

**Next:** [Mini Project](05_mini_project.qmd)
