---
title: "Welcome to the Workshop"
subtitle: "Getting Started with R"
---

## Learning Objectives

By the end of this module, you will:

1. Understand why R is valuable for biomedical research
2. Navigate the RStudio interface confidently
3. Know how this workshop is structured

## Why R?

- **Free and open source** - No license costs
- **Reproducible research** - Script everything
- **Powerful visualization** - Publication-quality figures
- **Active community** - Thousands of packages
- **Industry standard** - Used in pharma, biotech, academia

## RStudio Interface

```
+------------------+------------------+
|     Source       |   Environment    |
|    (Scripts)     |   (Variables)    |
+------------------+------------------+
|     Console      |      Files/      |
|   (Run code)     |  Plots/Help      |
+------------------+------------------+
```

### Key Panels

- **Source**: Write and save R scripts
- **Console**: Execute code interactively
- **Environment**: View your data and variables
- **Files/Plots/Help**: Navigate files, view plots, get help

## Your First R Commands

```{r}
# Basic math
2 + 2
10 / 3

# Create a variable
my_age <- 25
my_age

# Use a function
sqrt(16)
```

## Workshop Structure

| Day | Morning | Afternoon |
|-----|---------|-----------|
| 1 | R Basics, Data Import | Tidyverse, ggplot2 |
| 2 | Statistics, t-tests | ANOVA, Regression |
| 3 | Figures, Quarto | Mini-Project |

## Getting Help

```{r}
#| eval: false
?mean           # Help for a function
??regression    # Search help
```

---

**Next:** [R Primer](03_r_primer.qmd)
