---
title: "R Fundamentals"
subtitle: "Data Types and Structures"
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
```

## Learning Objectives

1. Understand R's basic data types
2. Create and manipulate vectors
3. Work with data frames
4. Use functions effectively

## Data Types

```{r}
# Numeric
age <- 25
height <- 1.75

# Character (text)
name <- "Dr. Mitra"
department <- "Bio Engineering"

# Logical
is_student <- TRUE
has_phd <- FALSE

# Check types
class(age)
class(name)
class(is_student)
```

## Vectors

Vectors are collections of values of the same type:

```{r}
# Numeric vector
ages <- c(22, 25, 31, 28, 24)
ages

# Character vector
treatments <- c("Control", "Drug_A", "Drug_B")
treatments

# Operations on vectors
mean(ages)
max(ages)
ages + 10
```

### Accessing Elements

```{r}
ages[1]          # First element
ages[c(1, 3)]    # First and third
ages[ages > 25]  # Filter by condition
```

## Data Frames

Data frames are tables with rows and columns:

```{r}
# Create a data frame
patients <- data.frame(
  id = c("P001", "P002", "P003"),
  age = c(45, 32, 58),
  treatment = c("Drug_A", "Control", "Drug_A"),
  response = c(TRUE, FALSE, TRUE)
)

patients
```

### Accessing Data

```{r}
patients$age           # Single column
patients[1, ]          # First row
patients[, "treatment"] # By column name
```

## Functions

```{r}
# Built-in functions
mean(ages)
sd(ages)
length(ages)
summary(ages)
```

### Getting Help

```{r}
#| eval: false
?mean
args(mean)
example(mean)
```

## Exercise

::: {.callout-caution title="Try It"}
1. Create a vector of 5 cell viability values (0-100)
2. Calculate the mean and standard deviation
3. Create a data frame with sample_id, treatment, and viability columns
:::

::: {.callout-tip collapse="true" title="Solution"}
```{r}
viability <- c(95, 87, 72, 91, 88)
mean(viability)
sd(viability)

experiment <- data.frame(
  sample_id = paste0("S", 1:5),
  treatment = c("Control", "Drug", "Drug", "Control", "Drug"),
  viability = viability
)
experiment
```
:::

---

**Next:** [Data Import](03_data_import.qmd)
