---
title: "Data Import"
subtitle: "Reading External Data into R"
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(here)
```

## Learning Objectives

1. Import CSV files with `read_csv()`
2. Use the `here` package for file paths
3. Handle common import issues
4. Explore imported data

## The here Package

```{r}
#| eval: false
library(here)

# Always works, regardless of working directory
here("data", "raw", "patients.csv")
```

## Reading CSV Files

```{r}
#| eval: false
library(tidyverse)

# Read a CSV file
data <- read_csv(here("data", "raw", "experiment.csv"))
```

### Common Options

```{r}
#| eval: false
read_csv(
  "file.csv",
  skip = 1,              # Skip header rows
  na = c("", "NA", "-"), # Define missing values
  col_types = cols(      # Specify column types
    id = col_character(),
    value = col_double()
  )
)
```

## Reading Excel Files

```{r}
#| eval: false
library(readxl)

# Read Excel file
data <- read_excel(
  here("data", "raw", "data.xlsx"),
  sheet = "Sheet1"
)

# List available sheets
excel_sheets(here("data", "raw", "data.xlsx"))
```

## Exploring Your Data

```{webr-r}
# Use built-in data for demonstration
data <- mtcars

# First look
head(data)
glimpse(data)

# Summary statistics
summary(data)

# Dimensions
print(paste("Rows:", nrow(data)))
print(paste("Cols:", ncol(data)))

# Column names
names(data)
```

## Common Issues

### Problem: Wrong delimiter

```{r}
#| eval: false
# For semicolon-separated files
read_csv2("file.csv")

# For tab-separated files
read_tsv("file.txt")
```

### Problem: Encoding issues

```{r}
#| eval: false
read_csv("file.csv", locale = locale(encoding = "UTF-8"))
```

## Exercise

::: {.callout-caution title="Try It"}
1. Load the `tidyverse` and `here` packages
2. Use `glimpse()` on the `iris` dataset
3. How many rows and columns does it have?
:::

::: {.callout-tip collapse="true" title="Solution"}
```{webr-r}
library(tidyverse)
glimpse(iris)
print(paste("Rows:", nrow(iris), "Cols:", ncol(iris)))
```
:::

---

**Next:** [Introduction to the Tidyverse](04_tidyverse_intro.qmd)
