---
title: "Introduction to the Tidyverse"
subtitle: "The Pipe Operator and Tidy Data"
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
```

## Learning Objectives

1. Understand the tidyverse philosophy
2. Master the pipe operator `|>`
3. Recognize tidy data principles

## What is the Tidyverse?

A collection of R packages designed for data science:

- **dplyr**: Data manipulation
- **ggplot2**: Visualization
- **tidyr**: Data tidying
- **readr**: Data import
- **purrr**: Functional programming

```{r}
library(tidyverse)  # Loads all core packages
```

## The Pipe Operator

The pipe `|>` passes the result of one function to the next:

```{r}
# Without pipe (nested functions)
round(mean(c(1, 2, 3, 4, 5)), 2)

# With pipe (left to right)
c(1, 2, 3, 4, 5) |>
  mean() |>
  round(2)
```

### Reading Pipes

Read `|>` as **"and then"**:

```{r}
mtcars |>
  filter(mpg > 20) |>    # Filter rows, and then
  select(mpg, cyl) |>    # Select columns, and then
  head(5)                 # Show first 5
```

## Tidy Data Principles

1. Each variable is a column
2. Each observation is a row
3. Each value is a cell

### Messy vs Tidy

```{r}
# Messy (wide format)
messy <- tibble(
  patient = c("A", "B"),
  day1 = c(100, 95),
  day2 = c(105, 98)
)
messy

# Tidy (long format)
tidy <- messy |>
  pivot_longer(
    cols = starts_with("day"),
    names_to = "day",
    values_to = "value"
  )
tidy
```

## Tibbles vs Data Frames

Tibbles are modern data frames:

```{r}
# Create a tibble
my_tibble <- tibble(
  x = 1:5,
  y = x^2,
  z = c("a", "b", "c", "d", "e")
)
my_tibble
```

## Exercise

::: {.callout-caution title="Try It"}
Use the pipe to:
1. Take the `iris` dataset
2. Filter for setosa species
3. Calculate mean Sepal.Length
:::

::: {.callout-tip collapse="true" title="Solution"}
```{r}
iris |>
  filter(Species == "setosa") |>
  summarize(mean_sepal = mean(Sepal.Length))
```
:::

---

**Next:** [Data Wrangling with dplyr](05_dplyr_wrangling.qmd)
