---
title: "Data Wrangling with dplyr"
subtitle: "The Five Core Verbs"
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
```

## Learning Objectives

1. Master the 5 core dplyr verbs
2. Chain operations with the pipe
3. Perform grouped summaries

## Sample Data

```{r}
set.seed(2025)
patients <- tibble(
  id = paste0("P", sprintf("%03d", 1:20)),
  age = round(rnorm(20, 55, 15)),
  treatment = rep(c("Drug_A", "Placebo"), each = 10),
  response = c(rnorm(10, 75, 10), rnorm(10, 60, 12))
)
head(patients)
```

## 1. filter() - Keep Rows

```{r}
# Single condition
patients |> filter(age > 60)

# Multiple conditions (AND)
patients |> filter(age > 50, treatment == "Drug_A")

# OR condition
patients |> filter(age < 40 | age > 70)
```

## 2. select() - Keep Columns

```{r}
# By name
patients |> select(id, response)

# Remove columns
patients |> select(-age)

# Helpers
patients |> select(starts_with("treat"))
```

## 3. mutate() - Create Columns

```{r}
patients |>
  mutate(
    response_pct = response / 100,
    age_group = if_else(age >= 60, "Senior", "Adult")
  )
```

## 4. arrange() - Sort Rows

```{r}
# Ascending
patients |> arrange(age)

# Descending
patients |> arrange(desc(response))

# Multiple columns
patients |> arrange(treatment, desc(response))
```

## 5. summarize() - Aggregate

```{r}
patients |>
  summarize(
    n = n(),
    mean_response = mean(response),
    sd_response = sd(response)
  )
```

## group_by() + summarize()

```{r}
patients |>
  group_by(treatment) |>
  summarize(
    n = n(),
    mean_response = mean(response),
    sd_response = sd(response)
  )
```

## Chaining It All Together

```{r}
patients |>
  filter(age >= 40) |>
  mutate(age_decade = floor(age / 10) * 10) |>
  group_by(treatment, age_decade) |>
  summarize(
    n = n(),
    mean_response = mean(response),
    .groups = "drop"
  ) |>
  arrange(desc(mean_response))
```

## Exercise

::: {.callout-caution title="Try It"}
Using `mtcars`:
1. Filter cars with mpg > 20
2. Create a new column `kpl` (mpg * 0.425)
3. Group by `cyl` and calculate mean kpl
:::

::: {.callout-tip collapse="true" title="Solution"}
```{r}
mtcars |>
  filter(mpg > 20) |>
  mutate(kpl = mpg * 0.425) |>
  group_by(cyl) |>
  summarize(mean_kpl = mean(kpl))
```
:::

---

**Next:** [ggplot2 Basics](06_ggplot2_basics.qmd)
