---
title: "ggplot2 Basics"
subtitle: "The Grammar of Graphics"
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
```

## Learning Objectives

1. Understand the grammar of graphics
2. Create common plot types
3. Customize plots with scales and themes

## The Grammar of Graphics

Every ggplot has three essential components:

1. **Data** - The dataset
2. **Aesthetics** - Map variables to visual properties
3. **Geoms** - Geometric objects to display

```{r}
ggplot(data = mtcars, aes(x = wt, y = mpg)) +
  geom_point()
```

## Common Geoms

### Scatter Plot

```{r}
ggplot(mtcars, aes(x = wt, y = mpg, color = factor(cyl))) +
  geom_point(size = 3)
```

### Box Plot

```{r}
ggplot(mtcars, aes(x = factor(cyl), y = mpg)) +
  geom_boxplot(fill = "steelblue", alpha = 0.7)
```

### Histogram

```{r}
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(bins = 10, fill = "steelblue", color = "white")
```

### Bar Plot

```{r}
ggplot(mtcars, aes(x = factor(cyl))) +
  geom_bar(fill = "steelblue")
```

### Line Plot

```{r}
# Create time series data
growth <- tibble(
  time = 0:10,
  od = 0.1 * exp(0.3 * time)
)

ggplot(growth, aes(x = time, y = od)) +
  geom_line(linewidth = 1, color = "steelblue") +
  geom_point(size = 2)
```

## Adding Labels

```{r}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  labs(
    title = "Fuel Efficiency vs Weight",
    x = "Weight (1000 lbs)",
    y = "Miles per Gallon"
  )
```

## Themes

```{r}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  labs(title = "Clean Look") +
  theme_classic()
```

## Colors

```{r}
ggplot(mtcars, aes(x = wt, y = mpg, color = factor(cyl))) +
  geom_point(size = 3) +
  scale_color_viridis_d() +
  theme_classic()
```

## Faceting

```{r}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  facet_wrap(~cyl) +
  theme_classic()
```

## Exercise

::: {.callout-caution title="Try It"}
Create a plot with `iris`:
1. Scatter plot: Sepal.Length vs Sepal.Width
2. Color by Species
3. Add title and axis labels
4. Use theme_classic()
:::

::: {.callout-tip collapse="true" title="Solution"}
```{r}
ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
  geom_point(size = 2) +
  labs(
    title = "Iris Sepal Dimensions",
    x = "Sepal Length (cm)",
    y = "Sepal Width (cm)"
  ) +
  scale_color_viridis_d() +
  theme_classic()
```
:::

---

**Day 1 Complete!** See you tomorrow for Statistics & Analysis.
