---
title: "Hypothesis Testing"
subtitle: "Making Statistical Decisions"
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
```

## Learning Objectives

1. Understand null and alternative hypotheses
2. Interpret p-values correctly
3. Know Type I and Type II errors

## The Logic of Hypothesis Testing

1. State null hypothesis (H₀) and alternative (H₁)
2. Collect data
3. Calculate test statistic and p-value
4. Make decision based on significance level (α)

## Null Hypothesis (H₀)

The "no effect" or "no difference" hypothesis.

- "The drug has no effect on cell viability"
- "There is no difference between groups"

## Alternative Hypothesis (H₁)

What we're trying to show:

- "The drug affects cell viability"
- "There is a difference between groups"

## P-Value

The probability of observing results as extreme as ours, **if H₀ were true**.

### P-Value is NOT:

- ❌ The probability that H₀ is true
- ❌ The probability of a false positive
- ❌ The size of the effect

### Decision Rule

| P-value | Common Interpretation |
|---------|----------------------|
| p < 0.001 | Very strong evidence |
| p < 0.01 | Strong evidence |
| p < 0.05 | Evidence |
| p ≥ 0.05 | Insufficient evidence |

## Types of Errors

|  | H₀ True | H₀ False |
|--|---------|----------|
| **Reject H₀** | Type I Error (α) | Correct |
| **Fail to Reject** | Correct | Type II Error (β) |

- **Type I**: False positive (seeing effect when none exists)
- **Type II**: False negative (missing a real effect)

## Effect Size Matters

Statistical significance ≠ practical significance

```{r}
# Tiny effect can be "significant" with large n
set.seed(42)
x <- rnorm(10000, 100, 15)
y <- rnorm(10000, 100.5, 15)  # 0.5 unit difference

t.test(x, y)
```

Small p-value, but difference of 0.5 may be meaningless.

## Best Practices

1. **State hypotheses before collecting data**
2. **Choose α before analysis** (usually 0.05)
3. **Report effect sizes, not just p-values**
4. **Consider practical significance**

---

**Next:** [t-Tests](03_ttests.qmd)
