---
title: "t-Tests"
subtitle: "Comparing Means"
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
```

## Learning Objectives

1. Perform one-sample, two-sample, and paired t-tests
2. Check assumptions
3. Interpret and report results

## Types of t-Tests

| Type | Use Case |
|------|----------|
| One-sample | Compare mean to known value |
| Two-sample | Compare means of two independent groups |
| Paired | Compare means of matched pairs |

## One-Sample t-Test

Is the mean different from a specific value?

```{r}
# Sample data
viability <- c(92, 88, 95, 91, 87, 93, 89, 94, 90, 86)

# Test: Is mean different from 90%?
t.test(viability, mu = 90)
```

## Two-Sample t-Test

Compare means of two independent groups:

```{r}
set.seed(2025)
control <- rnorm(20, 95, 5)
treatment <- rnorm(20, 85, 8)

t.test(treatment, control)
```

### Using Formula Syntax

```{r}
data <- tibble(
  group = rep(c("Control", "Treatment"), each = 20),
  value = c(control, treatment)
)

t.test(value ~ group, data = data)
```

## Paired t-Test

Same subjects measured twice (before/after):

```{r}
before <- c(120, 135, 128, 140, 132)
after <- c(115, 128, 122, 130, 126)

t.test(before, after, paired = TRUE)
```

## Checking Assumptions

### 1. Normality

```{r}
shapiro.test(viability)

# Visual check
qqnorm(viability)
qqline(viability)
```

### 2. Equal Variances (Two-sample)

```{r}
var.test(control, treatment)
```

If variances unequal, use Welch's t-test (the default in R).

## Reporting Results

> Viability was significantly lower in the treatment group (M = 85.2, SD = 7.8) compared to control (M = 95.1, SD = 4.9), t(38) = 4.73, p < .001.

## Exercise

::: {.callout-caution title="Try It"}
Test if mean mpg in `mtcars` is different from 20.
:::

::: {.callout-tip collapse="true" title="Solution"}
```{r}
t.test(mtcars$mpg, mu = 20)
```
:::

---

**Next:** [ANOVA](04_anova.qmd)
