---
title: "ANOVA"
subtitle: "Comparing Multiple Groups"
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
```

## Learning Objectives

1. Perform one-way ANOVA
2. Run post-hoc tests
3. Check assumptions

## When to Use ANOVA

- Comparing means across **3+ groups**
- t-test is for 2 groups; ANOVA extends to many

## Sample Data

```{r}
set.seed(2025)
experiment <- tibble(
  treatment = rep(c("Control", "Low", "Medium", "High"), each = 15),
  response = c(
    rnorm(15, 50, 8),
    rnorm(15, 55, 10),
    rnorm(15, 65, 9),
    rnorm(15, 75, 11)
  )
)
```

## One-Way ANOVA

```{r}
model <- aov(response ~ treatment, data = experiment)
summary(model)
```

## Interpretation

- **F value**: Ratio of between-group to within-group variance
- **Pr(>F)**: p-value for the overall test
- If p < 0.05: At least one group differs from others

## Post-Hoc Tests

ANOVA tells us *if* groups differ, not *which* ones. Use Tukey's HSD:

```{r}
TukeyHSD(model)
```

## Visualization

```{r}
ggplot(experiment, aes(x = treatment, y = response, fill = treatment)) +
  geom_boxplot(alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  scale_fill_viridis_d() +
  labs(title = "Response by Treatment", y = "Response") +
  theme_classic() +
  theme(legend.position = "none")
```

## Checking Assumptions

```{r}
# Normality of residuals
shapiro.test(residuals(model))

# Homogeneity of variance
bartlett.test(response ~ treatment, data = experiment)
```

## Exercise

::: {.callout-caution title="Try It"}
Test if Sepal.Length differs by Species in `iris`.
:::

::: {.callout-tip collapse="true" title="Solution"}
```{r}
model <- aov(Sepal.Length ~ Species, data = iris)
summary(model)
TukeyHSD(model)
```
:::

---

**Next:** [Correlation & Regression](05_correlation_regression.qmd)
