---
title: "Correlation & Regression"
subtitle: "Relationships Between Variables"
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(broom)
```

## Learning Objectives

1. Calculate and interpret correlation
2. Fit simple linear regression
3. Interpret regression output

## Correlation

Measures the strength and direction of a linear relationship.

```{r}
# Using mtcars
cor(mtcars$wt, mtcars$mpg)

# Correlation test
cor.test(mtcars$wt, mtcars$mpg)
```

### Interpretation

| r | Strength |
|---|----------|
| 0.0 - 0.3 | Weak |
| 0.3 - 0.7 | Moderate |
| 0.7 - 1.0 | Strong |

### Visualization

```{r}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "MPG vs Weight", x = "Weight (1000 lbs)", y = "MPG") +
  theme_classic()
```

## Simple Linear Regression

Predicting one variable from another:

$$y = \beta_0 + \beta_1 x + \epsilon$$

```{r}
model <- lm(mpg ~ wt, data = mtcars)
summary(model)
```

### Interpretation

- **Intercept**: Expected mpg when weight = 0
- **wt coefficient**: Change in mpg per unit increase in weight
- **R-squared**: Proportion of variance explained
- **p-value**: Significance of coefficient

## Tidy Output with broom

```{r}
library(broom)

# Coefficients
tidy(model)

# Model statistics
glance(model)
```

## Multiple Regression

```{r}
model2 <- lm(mpg ~ wt + hp + cyl, data = mtcars)
tidy(model2)
```

## Exercise

::: {.callout-caution title="Try It"}
1. Calculate correlation between Sepal.Length and Petal.Length in `iris`
2. Fit a regression model
3. What's the R-squared?
:::

::: {.callout-tip collapse="true" title="Solution"}
```{r}
cor(iris$Sepal.Length, iris$Petal.Length)

model <- lm(Petal.Length ~ Sepal.Length, data = iris)
summary(model)  # R-squared = 0.76
```
:::

---

**Next:** [Your Data Workshop](06_your_data_workshop.qmd)
