---
title: "Your Data Workshop"
subtitle: "Practice Session"
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
```

## Learning Objectives

1. Apply Day 1-2 skills to a dataset
2. Practice complete analysis workflow
3. Get help with your own data

## Workshop Datasets

Choose one of these datasets to work with:

### Option 1: mtcars
```{r}
glimpse(mtcars)
```

### Option 2: iris
```{r}
glimpse(iris)
```

### Option 3: Your Own Data
Bring your own CSV/Excel file!

## Suggested Workflow

### 1. Load and Explore
```{r}
#| eval: false
# Load data
data <- read_csv("your_file.csv")

# Explore
glimpse(data)
summary(data)
```

### 2. Descriptive Statistics
```{r}
#| eval: false
data |>
  group_by(group_variable) |>
  summarize(
    n = n(),
    mean = mean(outcome),
    sd = sd(outcome)
  )
```

### 3. Visualization
```{r}
#| eval: false
ggplot(data, aes(x = group, y = outcome)) +
  geom_boxplot() +
  theme_classic()
```

### 4. Statistical Test
```{r}
#| eval: false
# For 2 groups
t.test(outcome ~ group, data = data)

# For 3+ groups
aov(outcome ~ group, data = data) |> summary()
```

## Practice Exercise

Using `mtcars`:

1. Calculate mean mpg by number of cylinders
2. Create a boxplot of mpg by cylinders
3. Test if mpg differs by cylinders (ANOVA)
4. Run post-hoc test

::: {.callout-tip collapse="true" title="Solution"}
```{r}
# 1. Descriptive stats
mtcars |>
  group_by(cyl) |>
  summarize(mean_mpg = mean(mpg), sd_mpg = sd(mpg))

# 2. Visualization
ggplot(mtcars, aes(x = factor(cyl), y = mpg, fill = factor(cyl))) +
  geom_boxplot() +
  labs(x = "Cylinders", y = "MPG") +
  theme_classic() +
  theme(legend.position = "none")

# 3. ANOVA
model <- aov(mpg ~ factor(cyl), data = mtcars)
summary(model)

# 4. Post-hoc
TukeyHSD(model)
```
:::

---

**Day 2 Complete!** Tomorrow: Publication figures and Quarto reports.
