---
title: "Day 3 Exercises: Publication Figures"
subtitle: "Participant Version"
format: html
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(patchwork)
```

## Exercise 1: Publication Theme

### 1.1 Create a Custom Theme

Complete the theme function:

```{r}
#| eval: false
theme_pub <- function(base_size = 12) {
  theme_classic(base_size = base_size) +
    theme(
      plot.title = element_text(face = "___", hjust = 0),
      axis.title = element_text(face = "___"),
      panel.border = element_rect(fill = ___, color = "black"),
      legend.position = "___"
    )
}
```

### 1.2 Apply the Theme

Use your theme on a scatter plot:

```{r}
#| eval: false
ggplot(mtcars, aes(x = wt, y = mpg, color = factor(cyl))) +
  geom_point(size = 3) +
  labs(
    title = ___,
    x = ___,
    y = ___,
    color = "Cylinders"
  ) +
  ___  # Apply theme
```

---

## Exercise 2: Colorblind-Friendly Colors

### 2.1 Use Okabe-Ito Palette

```{r}
#| eval: false
okabe_ito <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442",
               "#0072B2", "#D55E00", "#CC79A7", "#999999")

ggplot(iris, aes(x = Species, y = Sepal.Length, fill = Species)) +
  geom_boxplot() +
  scale_fill_manual(values = ___) +
  theme_classic()
```

---

## Exercise 3: Multi-Panel Figures

### 3.1 Create Component Plots

```{r}
#| eval: false
# Plot A: Scatter plot
p1 <- ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  labs(title = "Weight vs MPG") +
  theme_classic()

# Plot B: Box plot (complete this)
p2 <- ggplot(mtcars, aes(___, ___)) +
  geom_boxplot() +
  labs(title = "MPG by Cylinders") +
  theme_classic()

# Plot C: Histogram (complete this)
p3 <- ggplot(mtcars, aes(___)) +
  geom_histogram(bins = 10, fill = "steelblue") +
  labs(title = "MPG Distribution") +
  theme_classic()
```

### 3.2 Combine Side by Side

```{r}
#| eval: false
___ + ___
```

### 3.3 Stack Vertically

```{r}
#| eval: false
___ / ___
```

### 3.4 Complex Layout

Create layout: (A + B) over C

```{r}
#| eval: false
(___ + ___) / ___
```

### 3.5 Add Panel Labels

```{r}
#| eval: false
(p1 + p2) / p3 +
  plot_annotation(tag_levels = ___)
```

---

## Exercise 4: Complete Figure

Create a publication-ready 2x2 figure for the iris dataset:

- Panel A: Sepal.Length vs Sepal.Width scatter (color by Species)
- Panel B: Petal.Length vs Petal.Width scatter (color by Species)
- Panel C: Sepal.Length boxplot by Species
- Panel D: Petal.Length boxplot by Species

Requirements:
- Use consistent theme
- Add panel labels (A, B, C, D)
- Collect legends
- Add figure title

```{r}
#| eval: false
#| fig-width: 10
#| fig-height: 8

# Panel A
fig_a <- ggplot(iris, aes(___)) +
  geom_point() +
  theme_classic()

# Panel B
fig_b <- ___

# Panel C
fig_c <- ___

# Panel D
fig_d <- ___

# Combine
(___) / (___) +
  plot_layout(guides = "collect") +
  plot_annotation(
    title = ___,
    tag_levels = "A"
  )
```

---

## Exercise 5: Export Figure

Complete the export code:

```{r}
#| eval: false
# Save as high-resolution TIFF
ggsave(
  filename = "___",
  width = ___,
  height = ___,
  dpi = ___
)

# Save as PDF
ggsave(
  filename = "___",
  width = ___,
  height = ___
)
```

---

## Bonus: Annotated Figure

Add statistical significance annotations to a bar plot:

```{r}
#| eval: false
# Summary data
summary_data <- mtcars |>
  group_by(cyl) |>
  summarize(
    mean_mpg = mean(mpg),
    se = sd(mpg) / sqrt(n())
  )

# Bar plot with error bars and significance
ggplot(summary_data, aes(x = factor(cyl), y = mean_mpg, fill = factor(cyl))) +
  geom_col() +
  geom_errorbar(aes(ymin = ___, ymax = ___), width = 0.2) +
  # Add annotation for significance (manually)
  annotate("text", x = 2, y = 28, label = "***") +
  labs(
    title = "Mean MPG by Cylinder Count",
    x = "Cylinders",
    y = "Mean MPG (± SE)"
  ) +
  theme_classic() +
  theme(legend.position = "none")
```
